### 2024-10-24 latex/paper_old.tex

\documentclass[12pt]{article}

\usepackage{sectsty}    % set section font
\usepackage{graphicx}
\usepackage{blindtext}  % generate random text
\usepackage{multicol}   % multiple columns
\usepackage[a4paper,
            bindingoffset=.2in,
            left=.45in,
            right=.65in,
            top=.8in,
            bottom=1in,
            footskip=.5in]  % page number
            {geometry}

% set font family
\usepackage{times}
\usepackage{anyfontsize}

% set roman numbers in sections
\renewcommand{\thesection}{\small\Roman{section}}
% set section and subsection font
\sectionfont{\fontsize{9}{1}\selectfont\sffamily}
\subsectionfont{\fontsize{9}{1}\selectfont\itshape}

% begin ---------------------------------------
\begin{document}

\fontsize{9}{10}\selectfont  % set font size and spacing

\begin{center}
% title
\section*{\huge{\sffamily Efficient string matching algorithms for large datasets}}
Giovanni Santini \\
\textit{Dept.\ of Information Engineering \\
and Computer Science} \\
\textit{University of Trento} \\
Tento, Italy \\
giovanni.santini@unitn.it \\
MAT. 325441
\end{center}

\bigskip

\begin{multicols}{2}

\textbf{\textit{Abstract}--
High-Performance Computing (HPC) technology is
impacting several industries, including the creative industries
and those operating in the Internet of Things. However, thus
far, scarce attention has been devoted by researchers to the use
of such technology in Internet of Sounds (IoS) settings. The
IoS is a new area emerging in industry and academia as an
extension of the IoT to the sonic domain, both in musical and
non-musical contexts. The IoS relates to a distributed network
of sound things, which are devices capable of sensing, acquiring,
processing, actuating, and exchanging data serving the purpose
of communicating sound-related information. In this position
paper, we investigate the integration of HPC technology with
the IoS. We illustrate such integration via a set of application
scenarios showing how the IoS can benefit from HPC.\ Finally, we
discuss the open challenges ahead of us to implement a successful
integration of these two kinds of technology. \\
\indent \textit{Index Terms}--Internet of Sounds, High-Performance
Computing, audio processing
}

\fontsize{9}{11}\selectfont  % set font size and spacing

\begingroup\centering\section{INTRODUCTION}\endgroup

% lorem ipsum
\blindtext\blindtext\blindtext

\begingroup\centering\section{RELATED WORK}\endgroup

\blindtext\blindtext\blindtext

\begingroup\centering\section{BODY}\endgroup

\blindtext\blindtext\blindtext

\begingroup\centering\section{EVALUATION}\endgroup

\blindtext\blindtext\blindtext

\begingroup\centering\section{CONCLUSIONS}\endgroup

\blindtext\blindtext\blindtext

\begingroup\centering\section{APPENDIX}\endgroup

\blindtext\blindtext\blindtext

\begingroup\centering\section{REFERENCES}\endgroup

\blindtext\blindtext\blindtext

% end -----------------------------------------
\end{multicols}
\end{document}