### 2024-11-24 latex/paper.tex
report: introduction

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{blindtext}  % generate random text
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Matrix Transposition and Symmetry: a case study on Superscalar Architectures\\
%{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
%should not be used}
%\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{Giovanni Santini}
\IEEEauthorblockA{\textit{Dept.\ of Information Engineering and Computer Science} \\
\textit{University of Trento}\\
Trento, Italy \\
giovanni.santini@unitn.it \\
MAT. 235441}
}

\maketitle

\begin{abstract}
TODO *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
parallel computing, matrix transpose, symmetrix matrix. benchmarking
\end{IEEEkeywords}

\section{Introduction}

Contemporary CPU architectures leverage superscalar designs to enhance performance, incorporating multiple cores and advanced Single Instruction Multiple Data (SIMD) extensions, such as Intel's Streaming SIMD Extensions (SSE) and Advanced Vector Extensions (AVX). Although SSE is not a new technology, being first introduced in the Pentium III in 1999 and subsequently expanded in later years \cite{b1}, modern algorithms leveraging SIMD instructions continue to be a topic of active research in recent literature.
For example, the FreeBSD's SIMD-enhanced libc is actively reworking the standard library's code. \cite{b2}.
In 2023, most of the string functions were reimplemented by Robert Clausecker et al on many architectures \cite{b3}, exploiting vector instructions and improving the overall performance of the standard library \cite{b4}.
Similarly, efficient multi-threaded algorithms are becoming more critical as the demand of computation increases worldwide due to AI training \cite{b5} \cite{b6}.
This case study examines the implementations and performance of two algorithms—out-of-place square matrix transposition (matTranspose) and square matrix symmetry checking (checkSymm)—that leverage superscalar architectures. Both algorithms are used pervasively in performance critical applications. For example, matrix transposition was used in NASA's Apollo 11 Guidance Computer (AGC) to convert vectors in platenary coordinate system to base reference system \cite{b7}. Additionally, matrix transposition is used by Fourier Transform in the West (FFTW) \cite{b8}, a popular library to calculate Fast Furier Transofrms, and in signal processing \cite{b9}. Furthermore, the current C++26 draft proposes basic linear algebra algorithms in the "linalg" header, including matrix transposition \cite{b10}.

In section 2, we discuss the state of the art of the two algorithms.
In section 3, we elaborate on the algorithms analyzed, discuss performance evaluation and we present a new benchmarking framework.
In section 4, we analyze experimental data and graphs.

\section{State of the Art}

- Review current research and developments related to your project \\
- Discuss existing solutions and their limitations \\
- Identift the gap your project aims to fill \\

\textit{We define the transpose of a matrix $A$, and we denote it by $A^t$, as the matrix obtained from $A$ by interchanging rows and columns. Specifically, if $A$ is an $m$-by-$n$ matrix, then $A^t$ is the $n$-by-$m$ matrix whose entries are given by the equation (1)}. \cite{b11}

\begin{equation}
	(A^t)_{k,j} = A_{j,k}
\end{equation}

Computing the transpose of a matrix is a well discussed problem.

Older fortran implementations were difficult due to fortran.

# a lot of In place and rectangular

Analysis of In-Situ Transposition
"Algorithms for in-place matrix transposition make use of the cyclic structure of
the transposition mapping. A clear introduction to the process was given in
1958 by Windley [6]. In 1968 Boothroyd presented ACM Algorithm 302 [1]. Then
in 1970 Laflin and Brefner presented ACM Algorithm 380 which they showed to
be significantly faster than Algorithm 302 [4]"

	[1] Transpose vector stored array.
	BOOTHROYD,Z. Algorithm 302: Transpose vector stored array. Go~r~m.A~M I0, ~ (1967),
	292-293

	anche 380 important, revised by 513

Naive algorithm does not exploit locality of reference. Recursive solutions were proposed.

## out-of-core - where I/O needs to be minimized

 The oldest context in which the spatial locality of transposition seems to have been studied is for out-of-core operation (by Alltop, 1975), where the matrix is too large to fit into main memory ("core"). 
 Reviews of out-of-core algorithms, especially as applied to parallel computing, can be found in e.g. Suh & Prasanna (2002) and Krishnamoorth et al. (2004).

 An Efficient Algorithm for
 Out-of-Core Matrix Transposition
 Jinwoo Suh, Member, IEEE, and Viktor K. Prasanna, Fellow, IEEE

 	"These efforts have focused on reducing the number
 	of I/O operations. However, in state-of-the-art architectures, memory-memory data transfer time and index computation time are also
 	significant components of the overall time. In this paper, we propose an algorithm that considers the index computation time and the
 	I/O time and reduces the overall execution time."

FFTW - MPI shit
/* Recursive "radix-r" distributed transpose, which breaks a transpose
over p processes into p/r transposes over r processes plus r
transposes over p/r processes.  If performed recursively, this
produces a total of O(p log p) messages vs. O(p^2) messages for a
direct approach.
*/

## GPUs

In-Place Transposition of Rectangular Matrices on Accelerators

An Efficient Matrix Transpose in CUDA C/C++
Feb 18, 2013
By Mark Harris

\textit{A square matrix $A$ is called symmetric if it equals its transpose} \cite{b100}.

\section{Contribution and Methodology}

- Elaborate on the unique contributions of your project \\
- Describe the methodology proposed, including algorithms (pseudo-code),
data structures, and parallelization techniques. \\
- Discuss the challenges faced and how they were addressed \\

\blindtext\blindtext

\section{Experiments and System Description}

- Detailed description of the computing system and platform. \\
- Relevant specifications or configurations (e.g., libraries
and programming toolchains). \\
- Description of the experimental setup, procedures, and
methodologies used in the project. \\
- Discussion on how experiments are designed to test the hypotheses
or achieve the objectives \\

\blindtext\blindtext

\section{Results and Discussion}

- Presentation of results \\
- Analysis and interpretation in context \\
- Comparison with the state-of-the-art \\

\blindtext\blindtext

\section{Conclusions}

- Summary of the key findings and contributions

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} Intel® 64 and IA-32 Architectures Software Developer’s Manual Combined Volumes: 1, 2A, 2B, 2C, 2D, 3A, 3B, 3C, 3D, and 4. Volume 1, Section 2.2.7 "SIMD Instructions"
\bibitem{b2} SIMD man page "https://man.freebsd.org/cgi/man.cgi?simd(7)"
\bibitem{b3} Git blame on freeBSD's source code, in "lib/libc/amd64/string", shows recent activity
\bibitem{b4} https://freebsdfoundation.org/blog/a-sneak-peek-simd-enhanced-string-functions-for-amd64/
\bibitem{b5} Amodei, D. Hernandez, D. AI and Compute, https://openai.com/blog/ai-and-compute
\bibitem{b6} Andrew J. Lohn, Micah Musse. AI and Compute How Much Longer Can Computing Power Drive Artificial Intelligence Progress? CSET Issue Brief
\bibitem{b7} Digitalized Apollo 11 guidance computer by Virtual AGC and MIT Museum. Code in "Apollo-11/Comanche055/PLANETERY\_INTERNAL\_ORIENTATION.agc" line 34; "Apollo-11/Luminary099/ATTITUDE\_MANEUVER\_ROUTINE.agc" line 441
\bibitem{b8} Matteo Frigo and Steven G. Johnson. "The design and implementation of FFTW3". The word "transpose" is mentioned 21 times
\bibitem{b9} M. Henriksson and O. Gustafsson, "Streaming Matrix Transposition on FPGAs Using Distributed Memories," 2023 IEEE Nordic Circuits and Systems Conference (NorCAS), Aalborg, Denmark, 2023, pp. 1-6, doi: 10.1109/NorCAS58970.2023.10305472.
\bibitem{b10} "https://www.open-std.org/jtc1/sc22/wg21/docs/papers/2023/p1673r12.html" Accessed in november 2024
\bibitem{b11} Sheldon Axler. Linear Algebra Done Right ourth edition
16 November 2024. Section 3C Matrices, definition 3.54
\bibitem{b12} Sheldon Axler. Linear Algebra Done Right ourth edition
16 November 2024. Section 9A Bilinear Forms and Quadratic Forms, definition 9.11

\end{thebibliography}
\vspace{12pt}

\end{document}