### 2024-12-10 latex/paper2.tex
mpi: setup

\iffalse

References draft:

Jaeyoung Choi, Jack J. Dongarra, David W. Walker.  Parallel Matrix Transpose Algorithms on Distributed Memory Concurrent Computers.  Mathematical Sciences Section Oak Ridge National Laboratory P.O. Box 2008, Bldg. 6012 Oak Ridge, TN 37831-6367

Notes: The communication is based of the greatest commond divisor (GCD) of the number of rows and columns (P and Q). If P and Q are relatively prima, the matrix transpose algorithm involves complete exchange communication. A block is LCM/P x LCM/Q if not relativaly prime and only LCM/GCD communications are required.

Jonathan Eckstein, Gyorgy Matyasfalvi. Efficient Distributed-Memory Parallel Matrix-Vector Multiplication with Wide or Tall Unstructured Sparse Matrices.

John C. Bowman, Malcolm Roberts. Adaptive Matrix Transpose Algorithms for Distributed Multicore Processors.

Notes: "The most popular algorithms for transposing an NxN matrix distributed over P processes are the direct communicatoin (all-to-all, simpler implementation) and recursive binary exchange (butterfly) algorithms."

Jaeyoung Choi, Jack Dongarra, David W. Walker. The design of scalable software libraries for distributed memory concurrent computers. Oak Ridge National Laboartory Mathematical Sciences Section. Oak RIdge, TN 37831-6367.
- they explained (apparently) block cyclic data distribution

So we learned that block cyclic data distribution (round robin) is efficient for data access. You can mix blocks with cycles to get a few combinations to try out.

\fi